---
title: "Week 3 Project 1"
author: "Hope (Amal) Jasentuliyana"
date: "2023-09-27"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

```

## Observations of csv datafile:
### Record-level (line by line)
* Every 4th line, starting at line 1 is dashes -----
* The first four lines [1-4] can be discarded (headers and dashes)
* Line 5 starts the data. Call this line n
* Line n + 1 contains the State & pre-rating
* Line n + 2 is dashes and can be discarded.
* Line n + 4 will be the start of the next record
* As above a record in a dataframe contains data from 2 lines from the csv

### first pass:

Assuming there are no abberant records, it looks like I can:
1. delete every 4th line
2. join every odd line with the successive line (delete carriage return)
3. separate on | and trim()  to get fields


```{r}
if (! require(readr)) {
  install.packages("readr")
  library(readr)
}


if (! require(stringr)) {
  install.packages("stringr")
  library(stringr)
}

if (! require(dplyr)) {
  install.packages("dplyr")
  library(dplyr)
}
if (! require(dplyr)) {
  install.packages("dplyr")
  library(dplyr)
}
df <- read_csv("./project1.txt", col_names = FALSE)
df

# I was going to use str_replace_all() to remove the dashes, but 
# since it's a dataframe, I can filter mod(3) != 1.


head(df)
tail(df)

#196 rows, which start and end on -----
# after filtering I should have (2 * 196 / 3 )- 1  rows

df2 <- df[1:nrow(df) %% 3 != 1, ]
# OR using dplyr
df3 <- df %>% filter(row_number() %% 3 != 1)

stopifnot(nrow(df2) != (2 * nrow(df)/3 -1))


#stopifnot(identical(df2,df3))  #assertion failed!
stopifnot(nrow(setdiff(df2,df3)) == 0)

#at this point, the dataframes have an equal number of rows, 
#and the no rows in df2 are not also in df3.
# since they aren't identical, the order must be different?




head(df3)

tail(df3)

```

At this point I am going to take a different approach and just do regex substitutions.

```{r}
myRegex = "^-+$"  # field consisting only of one or more dashes
df4 <- df[!grepl("^-+$",df$X1),]

stopifnot(nrow(df4)==nrow(df2))
stopifnot(identical(df2,df4))

View(df4)  #df2 an df4 have the same data. df3 may have  a different ordering.



```


### Field-level 

1. Field 1: Player/pair number [will be used to join with other players to get average pre-rating of opponents]
2. Field 2: Name  [text]
3. Field 3: Rating [decimal, 1 digit after decimal]
4. Field 4-10: {W,L,D}[:whitespace:]opponent_number. opponent_number is integer.
5. Field 11: State. [A-Z]{2}.  [text]
6. Field 12: [0-9]{8}\s+/R:\s+(\d){4}. pre-rating  [integer] 
7. Fields 13-20: unused for this analysis


