---
title: "Week 3 Project 1"
author: "Hope (Amal) Jasentuliyana"
date: "2023-09-27"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

```

## Observations of csv datafile:
### Record-level (line by line)
* Every 4th line, starting at line 1 is dashes -----
* The first four lines [1-4] can be discarded (headers and dashes)
* Line 5 starts the data. Call this line n
* Line n + 1 contains the State & pre-rating
* Line n + 2 is dashes and can be discarded.
* Line n + 4 will be the start of the next record
* As above a record in a dataframe contains data from 2 lines from the csv

### first pass:

Assuming there are no abberant records, it looks like I can:
1. delete every 4th line
2. join every odd line with the successive line (delete carriage return)
3. separate on | and trim()  to get fields


### Field-level 

1. Field 1: Player/pair number [will be used to join with other players to get average pre-rating of opponents]
2. Field 2: Name  [text]
3. Field 3: Rating [decimal, 1 digit after decimal]
4. Field 4-10: {W,L,D}[:whitespace:]opponent_number. opponent_number is integer.
5. Field 11: State. [A-Z]{2}.  [text]
6. Field 12: [0-9]{8}\s+/R:\s+(\d){4}. pre-rating  [integer] 
7. Fields 13-20: unused for this analysis

