---
title: "Data 607 Week3 Assignment"
author: "Hope (Amal) Jasentuliyana"
date: "2023-09-17"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Week 3


## Question 1.
```{r}
library(dplyr)
library(knitr)
library(stringr)
if (!requireNamespace("psych", quietly = TRUE)) {
  install.packages("psych")
}
library(psych)

dfMajors <-  read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv")
describe(dfMajors)
head(dfMajors)
```
there's one missing value in Major_Category
```{r}
dfMajors[is.na(dfMajors$Major_Category),]

```
## find majors that contain "DATA" or "STATISTICS"

```{r}

dfDataStatsMajors <- dfMajors[grep(".*(data|statistics).*",dfMajors$Major,ignore.case = TRUE),] # case insensitive search
dfDataStatsMajors

xd <- dfDataStatsMajors$Major
xd
```



### that works for two keywords, but would be cumbersome with more.
Here's one approach that might work better with thousands of keywords, and a large dataset but
I can't get it to work:


```{r}
if (!requireNamespace("stringi")) {
  install.packages("stringi")
}

library(stringi)

myKeywords = c("data","statistics")
dfMajors$lcMajor <- tolower(dfMajors$Major) # avoids overhead of case insensitive matching if we have many keywords
hasKeyword <- function(field, keywords) {
  any(stri_detect_fixed(field, keywords))
} #retruns a vector of logicals

majorsHaveKeywords <- sapply(dfMajors$lcMajor,hasKeyword, keywords=myKeywords )

dfDataStatsMajors2 <- dfMajors[majorsHaveKeywords,]
kable(dfDataStatsMajors2[,c("Major","Major_Category")], caption="Majors containing our keywords")  
describe(dfDataStatsMajors2)






```

## 2 Write code that transforms the data below:

### I don't understand the question.  I assume it's not asking for simple assignment like
myVector <- myDF$foods


### First  version I'm going to create a vector whose elements are strings which start and end with doublequotes
```{r}
foodVector  <- c("bell pepper"  ,"bilberry" ,    "blackberry"  , "blood orange",
 "blueberry"   , "cantaloupe" ,  "chili pepper" ,"cloudberry"  , "elderberry"  , "lime"    ,     "lychee"     ,  "mulberry"    ,
 "olive"      ,  "salal berry")
 #  make one flat string, each element gets a double quote prefix and suffix, and 
# each element is separated by commas
quotedFoodVector <- str_c('"',foodVector,'"')
quotedFoodVector
```
### second  i'm going to flatten the data into a single string that contains the source code to create the previous vector in R:
```{r}
myFoodString <- str_c('c(' , str_c(quotedFoodVector,collapse = ', '), ')')


writeLines(myFoodString)
#check it with the desired result:
myFoodString == 'c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")'
```


## 3 Describe, in words, what these expressions will match:

* (.)\1\1

#### If that was put into quotes and passed as a string to str_view("(.)\1\1"):

 The R parser will see \1 as ascii character code 001 and pass that as an argument to str_view90 so I believe it matches octal ascii code \001, "start of header" which I can't find documented too well: [https://www.ascii-code.com/character/%E2%90%81](https://www.ascii-code.com/character/%E2%90%81)

```{r}
# "\1" == "\001"
"\1" == "\001"
```

#### On the other hand, if "(.)\\1\\1 was passed to  str_view() and once in the function it sees (.)\1\1 (ie: I put a break point in str_view and (.)\1\1 is the value of the parameter):

this matches a character three times.  equivalent to (.){3}

* "(.)(.)\\2\\1"

This matches any two chars, followed by those characters in reverse order:
"abba",  
"1xx1",  
"#&&#"  
...etc

* (..)\1  

#### if that's passed in quoted like str_view("(..)\1")
again the R string parser will interpret the single backslash as octal \001

so this will match any two characters followed by \001

####  but if we pass it as "(..)\\1" and str_view() sees (..)\1

then it matches  a character, a second character, then the first character, then the second:  
abab
d1d1

* "(.).\\1.\\1"

Any character, a second character, the first character, a third character, and the first character again:
"a4aja"  
"88888"  
"&d&3&"  


* "(.)(.)(.).*\\3\\2\\1"  

any three characters (they can be different or the same), zero or more characters, followed by the third, second, and first matched chars:  
abccba
abcsdfsjsdfjlcba



## 4 Construct regular expressions to match words that:

* Start and end with the same character.
  * "(.).*\\1"  
* Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)
  * "(..).*\\1"
* Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)
  * "(.).*\\1.*\\1"