---
title: "CUNYSPS-Data607-week5"
author: "Hope Jasentuliyana"
date: "2023-10-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## dump of sample data from mySql
```sql
--
-- Table structure for table `flight_delays`
--

/* DROP TABLE IF EXISTS `flight_delays`; */
/*!40101 SET @saved_cs_client     = @@character_set_client */;
/*!50503 SET character_set_client = utf8mb4 */;
CREATE TABLE `flight_delays` (
  `idflight_delays` int NOT NULL AUTO_INCREMENT,
  `airline_name` varchar(45) NOT NULL,
  `delay_status` varchar(45) NOT NULL,
  `Los Angeles` int NOT NULL,
  `Phoenix` int NOT NULL,
  `San Diego` int NOT NULL,
  `San Franciso` int NOT NULL,
  `Seattle` int NOT NULL,
  PRIMARY KEY (`idflight_delays`),
  KEY `Airline_name` (`airline_name`),
  KEY `Delay_status` (`delay_status`)
) ENGINE=InnoDB AUTO_INCREMENT=5 DEFAULT CHARSET=utf8mb4 COLLATE=utf8mb4_0900_ai_ci;
/*!40101 SET character_set_client = @saved_cs_client */;

--
-- Dumping data for table `flight_delays`
--

LOCK TABLES `flight_delays` WRITE;
/*!40000 ALTER TABLE `flight_delays` DISABLE KEYS */;
INSERT INTO `flight_delays` VALUES (1,'ALASKA','on time',497,221,212,503,1841),(2,'ALASKA','delayed',62,12,20,102,305),(3,'AM WEST','on time',694,4840,383,320,201),(4,'AM WEST','delayed',117,415,65,129,61);
/*!40000 ALTER TABLE `flight_delays` ENABLE KEYS */;
UNLOCK TABLES;

```
## read data into a dataframe

```{r}
library(DBI)
library(dbplyr)
library(tidyverse)
library(RMariaDB)
library(glue)

# mysql user and pass set in environment vars
# use Sys.setenv() to manualy set user and password in current R session
db_user <- Sys.getenv("DB_USER")
db_pass <- Sys.getenv("DB_PASS")
con <- DBI::dbConnect(
  RMariaDB::MariaDB(),
  user = db_user,
  password = db_pass,
  host = "localhost",
  dbname = "DATA607_WEEK5"
)
table_name <- "flight_delays"
qry <- glue("SELECT * FROM {table_name}")
df <- dbGetQuery(conn = con, qry)
View(df)
```

## tidy data

All cols not in c(idflight_delays, airline_name, delay_status) need to be put into a long format with names_to="city", values_to="delay"

```{r}
df_long <- df %>%
  pivot_longer(
    cols = !c(idflight_delays, airline_name, delay_status),
    names_to = "city",
    values_to = "count_delayed",
    values_drop_na = TRUE
  )
```

## compare ALASKA and AM WEST
### group on airline and summarize sums

```{r}
df_summary <- df_long %>% 
  group_by(airline_name,delay_status) %>% 
  summarise(total_count=sum(count_delayed)) %>% 
  ungroup()
```

## compare airlines
### prepare contingency table
```{r}
contingency_table <- with(df_summary,xtabs(
  total_count ~ airline_name + delay_status
))

percent_delayed <-df_summary %>% 
  pivot_wider(names_from = delay_status,
              values_from = total_count) %>% 
  mutate(percent_delayed = delayed / (delayed + `on time`))
contingency_table
percent_delayed
```

13% of Alaska flights were delayed, compared to 11% of AM WEST flights.

### check chi-sq
```{r}

chisq_result <- chisq.test(contingency_table)
chisq_result
```
The differences are statistically significant.
